---
title: Exploring EU aquacultural economy (part_1)
author: Robert P. McGuinn
date: '2019-12-22'
slug: stecf-data-download-and-exploration
categories: []
tags: []
---

# purpose
 
Explore aquaculture economic data from the European Commisson, Fisheries and Aquaculture: Scientific, Technical and Economic Committee for Fisheries (STECF).

# analyst note 

This is part 1. As I move through the stages of exploration, and when I feel like it might be worth sharing with you, I'll post another part. Each part will include data and package load steps, so that they can be more stand-alone.

# data description 

Agency: European Commisson,Fisheries and Aquaculture.

https://stecf.jrc.ec.europa.eu/dd/aqua

# download data   

https://stecf.jrc.ec.europa.eu/documents/43805/2192243/STECF+18-19+-+Aquaculture+economic+data+tables.xlsx/bbbefd1b-62ca-45bd-a632-cfb97bd0e2ef

# original file format  

Excel spreadsheet(*.xlsx)
  
# original file name   

* 'STECF 18-19 - Aquaculture economic data tables.xlsx'    

Once you have the file downloaded, open it in Excel and take a look. I see that sheets 5 and sheets 6 are the 'DCF data' and the 'EUMAP data' respectively.  So go about uploading those two tables in R. 

# load packages  
The following list of packages are required to run the subsequent R commands. Install these if you do not already have them.

```{r load_packages, eval=TRUE, results='hide', message=FALSE, warning=FALSE}

library(openxlsx)
library(tidyverse)
library(magrittr)
library(flextable)

```

# load data to R

```{r load_data, eval=TRUE, cache=TRUE}

# load the data 
setwd("C:/rworking/fish.farmr/indata")

dcf <- read.xlsx("C:/rworking/fish.farmR/indata/STECF 18-19 - Aquaculture economic data tables.xlsx", sheet = 5, colNames = TRUE)

eumap <- read.xlsx("C:/rworking/fish.farmR/indata/STECF 18-19 - Aquaculture economic data tables.xlsx", sheet = 6, colNames = TRUE)

```
  
Now you have **'dcf'** and **'eumap'** data frames.  

# check the data frame size and variable names

Run the following exploratory commands.

```{r names_dcf, eval = TRUE}

# look at the variables and data fram dimensions for 'dcf'
names(dcf)
length(dcf)
dim(dcf)

# look at the variables and data frame dimensions for 'eumap'
names(eumap)
length(eumap)
dim(eumap)

# check for variable name mismatches between the two files
setdiff(names(dcf), names(eumap))
setdiff(names(eumap), names(dcf))

```

Running the commands above will give you a good idea of the data structure and size.

# now you can play!

First, generate some stats of interest.  These will then be available for placing in-line with your text descriptions of the data frames.  

```{r inline_calc_basic, eval = TRUE}

# number of records (dcf)
a <- length(dcf$country_code) 

# number of unique species(dcf)
b <- length(levels(factor(dcf$species)))

# number of records(eumap)
c <-  length(eumap$country_code)

# number of unique species(eumap)
d <- length(levels(factor(eumap$species)))

```


The number of records within **'dcf'** is `r prettyNum(a, big.mark = ",")`, with `r b` unique species, while the number of records within **'eumap'** are `r prettyNum(c, big.mark = ",")`, with `r d` unique 'species'.

# write inline code

The above paragraph uses inline code for the numerical values. Example below of what would be placed in an RMarkdown document, like this one.

```{r inline_example, eval = FALSE}

...number of records within dcf is `r prettyNum(a, big.mark = ",")`

```

# table: 'Total sales volume' by EU country

Here is an example summary table grouped by country with 'Total sales volume' in tonnes (all years together, from `r min(dcf$year)` to `r max(dcf$year)`. A range of `r max(dcf$year) - min(dcf$year)` years).

``` {r dcm_by_country, eval = TRUE}

x <- dcf %>% filter(value != 0, 
                    variable_name == 'Total sales volume'
                    # year == 2008
) %>% 
  group_by(country_name) %>%
  summarize(# species = paste(unique(species), collapse=" | "),
            variable = paste(unique(variable_name), collapse = " | "),
            # country_name = paste(unique(country_name), collapse = " | "),
            median_value = median(value),
            mean_value = mean(value),
            sum_of_value = sum(value),
            unit = paste(unique(unit), collapse=" | "),
            firms = n()) %>% arrange(-median_value) %>% 
  arrange(-sum_of_value)

#View(x)

# use 

colkeys <- names(x[,-1])

ft <- flextable(x) %>% 
  theme_vanilla() %>% #booktabs, vanilla, box, tron, tron_legacy
 # set_header_labels(Field_Names = "Field Name", n = "# of Records") %>% 
  colformat_num(col_keys = colkeys, digits = 0, big.mark = ",", na_str = "NA") %>% 
  set_header_labels(country_name = "country"
                    #species = "species_list"
                    ) %>% 
  font(fontname = "Cambria", part = "all") %>% #header, body, all
  fontsize(size = 14, part = "all") %>% 
  autofit() 

ft

```

# table: top 50 species by 'Sales volume production'

The following table gives the top 50 species with regards to 'Sales volume production' in unit = 'TONNE' within the EU. The **'dcf'** data was used here (all years together, from `r min(dcf$year)` to `r max(dcf$year)`). The sum of 'Sales volume production' grouped by species is used as the ranking variable.

``` {r eumap_by_country, eval = TRUE}

x <- dcf %>% filter(variable_name == 'Sales volume production', 
                    value != 0, 
                    is.na(value) == FALSE) %>% 
  
  group_by(scientific_name, common_name, species) %>%
  summarize(sum_value = sum(value)) %>% 
  arrange(-sum_value)


#View(x)

# Use the flextable package to make a nicely formatted table

x <- x[1:50,]

x$Rank <- seq_along(1:50)


colkeys <- names(x[,-1])

ft <- flextable(x) %>% 
  theme_vanilla() %>% #booktabs, vanilla, box, tron, tron_legacy
 # set_header_labels(Field_Names = "Field Name", n = "# of Records") %>% 
  colformat_num(col_keys = colkeys, digits = 0, big.mark = ",", na_str = "NA") %>% 
  # set_header_labels(#country_name = "country"
  #                   #species = "species_list"
  #                   ) %>% 
  font(fontname = "Cambria", part = "all") %>% #header, body, all
  fontsize(size = 14, part = "all") %>% 
  autofit() 

ft

```

# figure: quick plot of species ranks

Quick plot of species ranks to get a sense of the relative scales of production. 

```{r fig_species_rank_plot}

plot(x$sum_value)

```

It is evident that the top 2 species, `r x$common_name[1]` and `r x$common_name[2]`, are far and away the most produced by volume.

# analyst note
I think that is where I will leave this particular post. In part 2 of this exploration we will dig a little deeper with some nicer visualizations and maybe some more summary tables. For now you have the data in hand and can explore at will. Cheers!




