---
title: Exploring EU aquacultural economy (part_1)
author: Robert P. McGuinn
date: '2019-12-22'
slug: stecf-data-download-and-exploration
categories: []
tags: []
---

# purpose
 
Download and take an intial look at the European Commisson, Fisheries and Aquaculture: Scientific, Technical and Economic Committee for Fisheries (STECF) aquaculture economic data tables. 

# analyst note 

This is part 1. As I move through the stages of exploration, and when I feel like it might be worth sharing with you, I'll post another part. Each part will include data and package load steps, so that they can be more stand-alone.

# data description 

Agency: European Commisson,Fisheries and Aquaculture.

https://stecf.jrc.ec.europa.eu/dd/aqua

# download data   

https://stecf.jrc.ec.europa.eu/documents/43805/2192243/STECF+18-19+-+Aquaculture+economic+data+tables.xlsx/bbbefd1b-62ca-45bd-a632-cfb97bd0e2ef

# original file format  

Excel spreadsheet(*.xlsx)
  
# original file name   

* 'STECF 18-19 - Aquaculture economic data tables.xlsx'    

Once you have the file downloaded, open it in Excel and take a look. I see that sheets 5 and sheets 6 are the 'DCF data' and the 'EUMAP data' respectively.  So go about uploading those two tables in R. 

# load packages  
The following list of packages are required to run the subsequent R commands. Install these if you do not already have them.

```{r load_packages, eval=TRUE, results='hide', message=FALSE, warning=FALSE}

library(openxlsx)
library(tidyverse)
library(magrittr)
library(flextable)

```

# load data to R

```{r load_data, eval=TRUE, cache=TRUE}

# load the data 
setwd("C:/rworking/fish.farmr/indata")

dcf <- read.xlsx("C:/rworking/fish.farmR/indata/STECF 18-19 - Aquaculture economic data tables.xlsx", sheet = 5, colNames = TRUE)

eumap <- read.xlsx("C:/rworking/fish.farmR/indata/STECF 18-19 - Aquaculture economic data tables.xlsx", sheet = 6, colNames = TRUE)

```
  
Now you have **'dcf'** and **'eumap'** data frames.  

# check the data frame size and variable names

Run the following exploratory commands.

```{r names_dcf, eval = TRUE}

# look at the variables and data fram dimensions for 'dcf'
names(dcf)
length(dcf)
dim(dcf)

# look at the variables and data frame dimensions for 'eumap'
names(eumap)
length(eumap)
dim(eumap)

# check for variable name mismatches between the two files
setdiff(names(dcf), names(eumap))
setdiff(names(eumap), names(dcf))

```

Running the commands above will give you a good idea of the data structure and size.

# now you can play!

First, generate some stats of interest.  These will then be available for placing in-line with your text descriptions of the data frames.  

```{r inline_calc_basic, eval = TRUE}

# number of records (dcf)
a <- length(dcf$country_code) 

# number of unique species(dcf)
b <- length(levels(factor(dcf$species)))

# number of records(eumap)
c <-  length(eumap$country_code)

# number of unique species(eumap)
d <- length(levels(factor(eumap$species)))

```


The number of records within **'dcf'** is `r prettyNum(a, big.mark = ",")`, with `r b` unique species, while the number of records within **'eumap'** are `r prettyNum(c, big.mark = ",")`, with `r d` unique 'species'.

# write inline code

The above paragraph uses inline code for the numerical values. Example below of what would be placed in an RMarkdown document, like this one.

```{r inline_example, eval = FALSE}

...number of records within dcf is `r prettyNum(a, big.mark = ",")`

```

# table by country

Here is an example summary table grouped by country with total sales volume in tonnes (mean per firm value sum of all sales volume regardless of the year of reporting)

``` {r dcm_by_country, eval = TRUE}

x <- dcf %>% filter(value != 0, 
                    variable_name == 'Total sales volume'
                    # year == 2008
) %>% 
  group_by(country_name) %>%
  summarize(# species = paste(unique(species), collapse=" | "),
            variable = paste(unique(variable_name), collapse = " | "),
            # country_name = paste(unique(country_name), collapse = " | "),
            mean_value = mean(value),
            sum_of_value = sum(value),
            unit = paste(unique(unit), collapse=" | "),
            firms = n()) %>% arrange(-sum_of_value)

#View(x)

colkeys <- names(x[,-1])

ft <- flextable(x) %>% 
  theme_vanilla() %>% #booktabs, vanilla, box, tron, tron_legacy
 # set_header_labels(Field_Names = "Field Name", n = "# of Records") %>% 
  colformat_num(col_keys = colkeys, digits = 0, big.mark = ",", na_str = "NA") %>% 
  set_header_labels(country_name = "country"
                    #species = "species_list"
                    ) %>% 
  font(fontname = "Cambria", part = "all") %>% #header, body, all
  fontsize(size = 14, part = "all") %>% 
  autofit() 

ft

```










