---
title: STECF data download and inspect (part_1)
author: Robert P. McGuinn
date: '2019-12-21'
slug: stecf-data-download-and-exploration
categories: []
tags: []
---

# purpose
 
Download and take an intial look at the STECF aquaculture data in preparation for subsequent analysis and visualization.   

# analyst note 

This is part 1. As I move through the stages of exploration, and when I feel like it might be worth sharing with you, I'll post another part. 

# data description 

Agency: European Commisson: Fisheries and Aquaculture.

https://stecf.jrc.ec.europa.eu/dd/aqua

# download data   

https://stecf.jrc.ec.europa.eu/documents/43805/2192243/STECF+18-19+-+Aquaculture+economic+data+tables.xlsx/bbbefd1b-62ca-45bd-a632-cfb97bd0e2ef

# original file format  

Excel spreadsheet(*.xlsx)
  
# original file name   

* 'STECF 18-19 - Aquaculture economic data tables.xlsx'    

Once you have the file downloaded, open it in Excel and take a look. I see that sheets 5 and sheets 6 are the 'DCF data' and the 'EUMAP data' respectively.  So go about uploading those two tables in R. 

# load packages  
The following list of packages are required to run the subsequent R commands. Install these if you do not already have them.

```{r load_packages, eval=FALSE}

library(openxlsx)
library(tidyverse)
library(magrittr)

```

# load data to R

```{r load_data, eval=TRUE, cache=TRUE}

# load the data 
setwd("C:/rworking/fish.farmr/indata")

dcf <- read.xlsx("C:/rworking/fish.farmR/indata/STECF 18-19 - Aquaculture economic data tables.xlsx", sheet = 5, colNames = TRUE)

eumap <- read.xlsx("C:/rworking/fish.farmR/indata/STECF 18-19 - Aquaculture economic data tables.xlsx", sheet = 6, colNames = TRUE)

```
  
Now you have **'dcf'** and **'eumap'** data frames.  

# check the data frame size and variable names

Run the following exploratory commands.

```{r names_dcf, eval = TRUE}

# look and variables and df size
names(dcf)
length(dcf)
dim(dcf)

# look and variables and df size
names(eumap)
length(eumap)
dim(eumap)

# check for variable name mismatches
setdiff(names(dcf), names(eumap))
setdiff(names(eumap), names(dcf))

```

Running the commands above will give you a good idea of the data structure and size.

# now you can play!

First, generate some stats of interest.  These will then be available for placing in-line with your text descriptions of the data frames.  

```{r inline_calc_basic, eval = TRUE}

# number of records (dcf)
a <- length(dcf$country_code) 

# number of unique species(dcf)
b <- length(levels(factor(dcf$species)))

# number of records(eumap)
c <-  length(eumap$country_code)

# number of unique species(eumap)
d <- length(levels(factor(eumap$species)))

```


The number of records within **'dcf'** is `r prettyNum(a, big.mark = ",")`, with `r b` unique species, while the number of records within **'eumap'** are `r prettyNum(c, big.mark = ",")`, with `r d` unique 'species'.

# write inline code

The above paragraph uses inline code for the numerical values. Example below of what would be placed in an RMarkdown document, like this one.

```{r inline, eval = FALSE}

...number of records within dcf is `r prettyNum(a, big.mark = ",")`

```





